### 2018-02-28 dissertation/1-intro/chapter1.tex
dissertation v 2017-11-30
several updates to text; possible problem with character encoding

\section{Document organization}
The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
Finally, \autoref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.

>>>>

\section{Document organization}
The remainder of this dissertation is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
\Cref{chap:truman} describes the goals of TruMan, its underlying hypothesis and the work done to achieve those goals, as well as the previously existing algorithms which form parts of TruMan.
Next, \autoref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
Finally \autoref{chap:conclusion} presents the final thoughts on the project and concludes this document.

%Finally, \autoref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.

### 2018-02-28 dissertation/2-complexnetworks/chapter2.tex
several updates to text; possible problem with character encoding

Ad-hoc networks require a decentralized approach to trust management; each member of a network has its own opinions about other members, and these opinions can change over time.
For these opinions to be generated and updated, two nodes must have had previous contact with each other, or derive trust from a third, intermediary, node.
Hence, there is the correlation between the trust graph and the topology graph of the network.
Since MANETs are dynamic, the graph that represents one network's topology is frequently changing and, with that, the opportunities to create and update trust relationships also changes.
%For two nodes to have a trust relationship at all, the topology must changed in a way so that they are neighbors to each other, or close enough that asking for a third node's
In networks in which nodes can meet more than once, it might be valuable to store information from previous encounters to use in the future, although this process can be too slow or resource-consuming to be viable in certain devices; by doing this, the trust graph maintains edges between nodes that are no longer connected in the topology.

>>>>

Ad-hoc networks require a decentralized approach to trust management; each member of a network has its own opinions about other members, and these opinions can change over time.
For these opinions to be generated and updated, two nodes must have had previous contact with each other, or derive trust from a third, intermediary, node.
Hence, there is the correlation between the trust graph and the topology graph of the network.
Since MANETs are dynamic, the graph that represents a network's topology is frequently changing and, with that, the opportunities to create and update trust relationships also changes.
%For two nodes to have a trust relationship at all, the topology must changed in a way so that they are neighbors to each other, or close enough that asking for a third node's
In networks in which nodes can meet more than once, it might be valuable to store information from previous encounters to use in the future, although this process can be too slow or resource-consuming to be viable in certain devices; by doing this, the trust graph maintains edges between nodes that are no longer connected in the topology.

### 2018-02-28 dissertation/3-vanets/chapter3.tex
several updates to text; possible problem with character encoding

\section{Trust in VANETs}
\label{section:trustvanet}
Like in other types of networks, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) which participate in it.
If one node is malicious or faulty, it can spread incorrect data that may compromise the network's utility.
Once the concept of VANETs was established, researchers have been attempting to predict ways in which malicious users might use the network to their advantage.
Examples include triggering false alarms about inexistent accidents, lying about the average speed in one road to make it less desirable for others, and falsifying geolocation data to exploit location-based routing algorithms. Therefore, the concept of trust must be established in the vehicular network context, allowing for nodes to judge the validity of information transmitted by others and share those conclusions with other nodes.

@@

\begin{table}[t!]
\caption{Properties of Truman and related work}
\label{table:properties}
\centering
\begin{tabular}{|p{4cm}||p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|}
 \hline
 \textbf{Property} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 TruMan & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{dotzer2005vars} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{minhas2010towards} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{chen2010trust} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{park2011long} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{huang2014social} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{li2016art} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{chen2017cloud} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
\end{tabular}
\end{table}

@@

\cite{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node appends its opinion of the message's contents.
That opinion is formed using a combination of the forwarding node's own observations of the event, its opinion of the origin node and previous opinions appended to the message.
This process adds credibility to a message through validation by nodes in a decentralized fashion.
It combines aspect of data- and entity- based trust, since nodes share their opinion of the data as well as their opinion of the sender. 
One interesting observation is setting higher trust values for certain vehicles based on their familiarity with the region (vehicles that reside in a given city may have more experience with certain types of events than newcomers).
However, opinion piggybacking has its own share of problems.
First, it allows forwarding nodes to access (at least some of) the contents of a message so it can form an opinion on it, diminishing privacy; a malicious forwarding node could even attempt to alter those contents.
Second, using previously appended opinions from other nodes to form a new opinion causes the first nodes to forward the message to have a substantially greater impact over the final opinion than the later ones.
Finally, there is an issue with scalability, since appending new information to a message on each hop may add a significant overhead to the transmission. Additionally, the authors provide little to no experimentation or proof that their approach would be sound in a real-world network.

>>>>

\section{Trust in VANETs}
\label{section:trustvanet}
Like in other types of networks, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) which participate in it.
If a node is malicious or faulty, it can spread incorrect data that may compromise the network's utility.
Once the concept of VANETs was established, researchers have been attempting to predict ways in which malicious users might use the network to their advantage.
Examples include triggering false alarms about inexistent accidents, lying about the average speed in a road to make it less desirable for others, and falsifying geolocation data to exploit location-based routing algorithms.
Therefore, the concept of trust must be established in the vehicular network context, allowing for nodes to judge the validity of information transmitted by others and share those conclusions with other nodes.

@@

\begin{table}[t!]
\caption{Properties of Truman and related work}
\label{table:properties}
\centering
\begin{tabular}{|p{4cm}||p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|}
 \hline
 \textbf{Property} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 TruMan & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & - & \checkmark \\
% \hline
% \cite{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{dotzer2005vars} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{minhas2010towards} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -\\
 \hline
 \cite{chen2010trust} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & -\\
 \hline
 \cite{park2011long} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{huang2014social} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 \hline
 \cite{li2016art} & \checkmark & - & - & - & \checkmark & - & - & \checkmark\\
 \hline
 \cite{chen2017cloud} & - & \checkmark & - & \checkmark & - & \checkmark & \checkmark & -\\
 \hline
\end{tabular}
\end{table}

@@

\cite{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node appends its opinion of the message's contents.
That opinion is formed using a combination of the forwarding node's own observations of the event, its opinion of the origin node and previous opinions appended to the message.
This process adds credibility to a message through validation by nodes in a decentralized fashion.
It combines aspect of data- and entity- based trust, since nodes share their opinion of the data as well as their opinion of the sender. 
An interesting observation is setting higher trust values for certain vehicles based on their familiarity with the region (vehicles that reside in a given city may have more experience with certain types of events than newcomers).
However, opinion piggybacking has its own share of problems.
First, it allows forwarding nodes to access (at least some of) the contents of a message so it can form an opinion on it, diminishing privacy; a malicious forwarding node could even attempt to alter those contents.
Second, using previously appended opinions from other nodes to form a new opinion causes the first nodes to forward the message to have a substantially greater impact over the final opinion than the later ones.
Finally, there is an issue with scalability, since appending new information to a message on each hop may add a significant overhead to the transmission. Additionally, the authors provide little to no experimentation or proof that their approach would be sound in a real-world network.

### 2018-02-28 dissertation/4-project/chapter4.tex
several updates to text; possible problem with character encoding

This work proposes that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
%The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used as input for a malicious node detection algorithm.

@@

In this chapter, the details for the above process are presented.
First, it is shown how vehicles can form long-term relationships and trust one another in a similar way to social networks.
Then, the Working Day Movement Model is introduced, along with the reasons why it is best fitted for this work.
Next, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment, followed by a list of changes that must be made to the model in order for it to handle dynamic networks.
Finally, a schedule for the next steps of the project is shown.

\section{Goals}
\label{section:goals}

@@

The basis of the presented trust model is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, \cite{vernize2013dissertation}.
This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
Furthermore, the algorithm is executed by a global observer which has information about the complete network.

The input graph $T = (V,O)$ is a static, connected, and directed graph containing all trust relationships in the network.
Such relationships are binary, so there are no varying degrees of trust: either one node trusts another completely (edge value is $1$), or it distrusts the other completely (edge value is $0$).
The relationships are also directed, meaning that if the value of $A\rightarrow B$ is $1$, $B\rightarrow A$ is not necessarily $1$.

First, $T$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
In each of these components, all nodes are connected by edges of value $1$.
In other words, within a single component, all nodes trust one another; nodes connected by edges of value $0$ are separated into different components.
Each of these components becomes a node of a component graph $C = (V', O')$.

@@

@@

>>>>

This work shows that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
%The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used as in conjunction with other algorithms in order to classify nodes as correct or malicious.

@@

In this chapter, the reasoning behind TruMan and details of how it works are presented.
First, it is shown how vehicles can form long-term relationships and trust one another in a similar way to social networks.
Then, two algorithms are introduced: Tarjan's strongly connected components algorithm \cite{tarjan1972depth} and an efficient algorithm for graph coloring \cite{mittal2011graph}.
Next, the Malicious Node Identification Algorithm (MaNI) \cite{vernize2013dissertation} is explained, for it is the work that suggests the usage of strongly connected components and graph coloring for malicious node detection.
Finally, TruMan itself is explained, showing how it combines features from existing algorithms and adapts them to a dynamic environment, enabled by the social properties found in vehicular networks.
%Then, the Working Day Movement Model is introduced, along with the reasons why it is best fitted for this work.
%Next, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment, followed by a list of changes that must be made to the model in order for it to handle dynamic networks.
%Finally, a schedule for the next steps of the project is shown.

\section{Goals}
\label{section:goals}
Building from the foundation set by MaNI \cite{vernize2013dissertation}, TruMan strives to enable efficient trust management in highly dynamic networks such as VANETs.
In addition to efficiency, it is desirable that the trust model is both simple to understand and to implement, making it appealing for real-world usage.

Furthermore, the desired properties of a VANET trust model described in \cite{zhang2011survey} (explained in detail in \autoref{section:properties}) were considered, so TruMan attempts to fulfill or enable as many of those properties as possible.

@@

@@

	\If{weight of $u\rightarrow v < h$}
		\State \textbf{continue}
	\EndIf

@@

The basis of TruMan is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, which suggests the use of strongly connected components and graph coloring for malicious node detection.
This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
Furthermore, the algorithm is executed by a global observer which has information about the complete network.

The input graph $T = (V,E)$ is a static, connected, and directed graph containing all trust relationships in the network.
Such relationships are binary, so there are no varying degrees of trust: either one node trusts another completely (edge value is $1$), or it distrusts the other completely (edge value is $0$).
The relationships are also directed, meaning that if the value of $A\rightarrow B$ is $1$, $B\rightarrow A$ is not necessarily $1$.

First, $T$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
In each of these components, all nodes are connected by edges of value $1$.
In other words, within a single component, all nodes trust one another; nodes connected by edges of value $0$ are separated into different components.
Each of these components becomes a node of a component graph $C = (V', E')$.
